Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-02-15 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex
went through bounded examples

As observed for the linear shear flow, our simulations 
%with $0<\beta\le 10^{-2}$ 
%%in the range  $10^{-4}\le \beta\le 10^{-2}$ 
%\todo[inline]{I don't understand this range of $\beta$ values?}
show that the equilibrium reduced area of a semipermeable inextensible membrane in a
Poiseuille flow is independent of the initial reduced area.
%and $\beta$. 
 Instead we find that larger maximum flow velocity $U$ results
in smaller equilibrium reduced area
(Figure~\ref{fig:parabolicComposite}(a)). At the smallest velocity $U=
200\; \mu$m/s, the steady state reduced area is large, and the
equilibrium vesicle shape is an axisymmetric bullet. The equilibrium
reduced area decreases with increasing $U$, and the nonlinear flow
profile gives rise to an asymmetric tank-treading slipper for $U\ge 400$
$\mu m/s$. We observe that the membrane permeability to water
drastically alters the equilibrium vesicle shape and position relative
to the Poiseuille flow: At $\beta=10^{-3}$ the equilibrium vesicle
depends only on the maximum velocity $U$ of the nonlinear shear flow.
The higher $U$ the smaller the equilibrium reduced area and the farther
away the vesicle is relative to the flow center at $y=0$.

@@

\section{Hydrodynamics of A semipermeable inextensible membrane under strong confinement\label{sec:confinement}}
Here we consider a semipermeable inextensible membrane under strong
confinement: a long closely-fit channel and a contracting channel. In a
narrow microfluidic channel, a large pressure jump is often used to push
the vesicle through a closely-fit channel~\cite{abk-fai-sto2006}. Our
simulations show that such a configuration can lead to an amplification
of water permeation at a fast time scale of the order of
\todo[inline]{mini seconds?} mini seconds.  In addition, we find that a
large value of $\beta$ is needed to keep the membrane tension below the
threshold value for membrane poration. Such high permeability for water
may be regarded as an indication that the lipid bilayer membrane is
porated under strong confinement, and our model inextensible membrane is
for a porated vesicle within the Helfrich free energy framework. As in
the previous examples, we fix the membrane length and examine the
enclosed water content as a dynamic consequence of water flow, membrane
shape deformation, and membrane tension distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Closely-fit Channel (Stenosis)}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semipermeable inextensible membrane
  passing through a closely-fit channel similar to the experimental
  device in~\citet{abk-fai-sto2006}. (b) A semipermeable inextensible membrane passing
  through a (four-times) longer closely-fit channel than (a). In (a) and
  (b), red regions correspond to influx and blue regions correspond to
  efflux. (c) and (d) The tension in N/m of the vesicle shapes in (a)
  and (b). (e) The excess pressure of an impermeable and semipermeable
  inextensible membrane in the short geometry in (a). (f) The reduced area in the geometries in (a) and (b) as a function
  of the membranes' center of mass. In (e) and (f), the x-axis is scaled
  by $1/L$ where $L$ is half the length of the constriction. (g) The
  reduced area in the geometries in (a)
  and (b) as a function of time. The time axis is scaled by $1/t_R$
  where $t_R$ is the residency time. In (e) and (f), the dashed lines
  correspond to the locations of the inlet and outlet. In (g), the
  dashed lines correspond to the times that the vesicle enters and exits
  the constriction. The colored membranes show the tension distribution
  in N/m of the circled cases. Throughout the entire simulation,
  these membranes have the largest tension.}
\end{figure*}
When red blood cells go through small capillary vessels, they experience
large shape deformation and significant membrane stretching that
triggers ATP release~\cite{Wan2008_PNAS, ForsythWan2011_PNAS}. Many
cells exhibit extraordinary flexibility as they go through narrow
vessels~\cite{AuStoreyMoore2016_PNAS}, and adjustment of
surface-to-volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going
through a stenosis (a long microfluidic channel with a width smaller
than the cell radius). Based on previous
findings~\cite{abk-fai-sto2006}, we first examine the effect of
semipermeability on the pressure fluctuation across a stenosed geometry.
In this experiment, a pressure drop is used to squeeze a red blood cell
through a constriction with diameter $5\;\mu$m and length $45\;\mu$m. We
set up a similar computational geometry with a Poiseuille flow imposed
at the inlet and outlet (Figure~\ref{fig:stenosisComposite}(a)). We set
the maximum flow velocity to $U_{\max} = 1000 \;\mu$m/s at the inlet and
outlet so that the residency time of the membrane in the stenosis channel
is $t_R \approx 25$ ms which agrees with the experimental
results~\cite{abk-fai-sto2006}. When the red blood cell enters the
channel, an increase in the pressure drop is required to maintain a
constant flow rate, and the difference in the pressure drop is called
the excess pressure.

@@

Figures~\ref{fig:stenosisComposite}(f)
and~\ref{fig:stenosisComposite}(g) show that the semipermeable inextensible membrane deflates from
the beginning due to the high flow velocity, consistent with the results
for a vesicle in a Poiseuille flow in open space. As the membrane enters
the stenosis channel, it continues to deflate in the front even though
we observe influx (red segment) at various rear locations in the
membrane. Overall, the membrane loses about 6.1\% of its initial total
water at a constant rate as shown in
Figures~\ref{fig:stenosisComposite}(f)
and~\ref{fig:stenosisComposite}(g) (blue curves). Thus we expect that a
larger amount of deflation is possible if the inextensible membrane is inside a longer
stenosis: we consider a vesicle passing through a channel of triple the
length in Figure~\ref{fig:stenosisComposite}(b) with
$U_{\max}=1000\;\mu$m/s, and we observe that vesicle stays inside the
stenosis for a residence time $t_R \approx 70$ ms and the area decreases
by 13.2\% (red curves).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Contracting Geometry} 
\begin{figure}[htp]
  \centering
    \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting channel
  similar to the experimental configuration in~\citet{wu2015critical}.
  The channel width varies from $18\;\mu$m to $1.8\;\mu$m, and it
  repeats every $45\;\mu$m. The fluid flux along the red regions is into
  the vesicle, and the fluid flux along the blue regions is out of the
  vesicle.  (a--c) The ability for a semipermeable vesicle to pass
  through the contraction depends on the permeability $\beta$. (d--f)
  The ability for a semipermeable inextensible membrane to pass through the
  contraction depends on the flow rate. (g--i) The tension distribution
  and maximum tension in N/m of the vesicle shapes in (a) and (b).
  (j--l) The tension distribution and maximum tension in N/m of the
  vesicle shapes in (d) and (e).}
\end{figure}

Squeezing a cell or a vesicle through a slit in microfluidics
has gained popularity for its application in stress-induced release of
macromolecules~\cite{ShareiEtAl2013_PNAS, Pak2015_PNAS,
ZhangShenHoganBarakatMisbah2018_BJ, LuoBai2019_PoF}. Red blood cells go
through submicron slits many times during their lifetime, and this
process selects healthy, flexible red blood cells to
survive~\cite{wu2015critical,LuPeng2019_PoF}. Here we investigate the
role of membrane permeability to water when a semipermeable inextensible membrane is
squeezed through a contracting microfluidic channel similar
to~\citet{wu2015critical}, where the width of the channel gradually
decreases from $18 \;\mu$m to $1.8\; \mu$m. The channel then immediately
reopens to $18\;\mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant
maximum flow rate of $U_{\max}$. We initialize a semipermeable inextensible membrane
with $\alpha = 0.65$ that spans $10\; \mu$m of the total channel width
and investigate the effects of the membrane permeability to water and
the maximum flow velocity on whether a membrane can successfully pass
through the slit. We color the vesicle according to the sign of its
flux---red for influx and blue for efflux.

@@

\begin{figure}[hbp]
  \centering
  \ifTikz
  \input{figures/contractingComposite2.tikz}
  \else
  \includegraphics{figures/contractingComposite2.pdf}
  \fi
  \caption{\label{fig:contractingComposite2} Reduced area and
  velocity of a semipermeable inextensible membrane vs.~time. The permeability coefficient is $\beta = 10^{-3}$
  and the maximum velocity is $18\;\mu$m/s. (a) A semipermeable inextensible membrane
  passing through six periods of the contracting geometry. (b) The
  reduced area with the marks representing shapes in (a). (c) The
  center-of-mass velocity with the marks representing shapes in (a). In (b) and
  (c), the lower $x$-axis is the x-coordinate of the center of mass, and
  the upper $x$-axis is time. (d) The membrane tension in N/m. (e) The
  maximum membrane tension.}
%  The colored vesicles show the tension
%  distribution in N/m of the circled vesicles. Throughout the entire
%  simulation, these vesicle configurations have the largest tension.}
\end{figure}
To understand the effects of water permeability on a semipermeable inextensible membrane that
repeatedly passes through micro-capillary vessels or sub-micron slits,
we construct a contracting geometry
(Figure~\ref{fig:contractingComposite2}) to simulate a membrane passing
through the contracting geometry six times. In
Figure~\ref{fig:contractingComposite2}(b), we show that the membrane
deflates each time it passes through the contraction followed by a
gradual inflation (over seconds). The total area loss over the history
of the vesicle is almost 50\%. In
Figure~\ref{fig:contractingComposite2}(c), we show the membrane center-of-mass velocity
as a function of the membrane position. As the membrane passes
through the contraction, it first accelerates until it reaches the neck
where it abruptly decelerates to go through the contraction.
Figure~\ref{fig:contractingComposite2}(d) shows the tension
distribution, and the maximum tension is in
Figure~\ref{fig:contractingComposite2}(e). We see that the membrane
tensions is largest when it passes through the neck, and it is
particularly large when it passes through the first neck since this is
when the membrane has the largest area.

>>>>

As observed for the linear shear flow, our simulations show that the
equilibrium reduced area of a semipermeable inextensible membrane in a
Poiseuille flow is independent of the initial reduced area. Instead we
find that larger maximum flow velocity $U$ results in smaller
equilibrium reduced area (Figure~\ref{fig:parabolicComposite}(a)). At
the smallest velocity $U= 200\; \mu$m/s, the steady state reduced area
is large, and the equilibrium vesicle shape is an axisymmetric bullet.
The equilibrium reduced area decreases with increasing $U$, and the
nonlinear flow profile gives rise to an asymmetric tank-treading slipper
for $U\ge 400$ $\mu m/s$. We observe that the membrane permeability to
water drastically alters the equilibrium vesicle shape and position
relative to the Poiseuille flow: At $\beta=10^{-3}$ the equilibrium
vesicle depends only on the maximum velocity $U$ of the nonlinear shear
flow. The higher $U$ the smaller the equilibrium reduced area and the
farther away the vesicle is relative to the flow center at $y=0$.

@@

\section{Hydrodynamics of A semipermeable inextensible membrane under strong confinement\label{sec:confinement}}
Here we consider a semipermeable inextensible membrane under strong
confinement: a long closely-fit channel and a contracting channel. In a
narrow microfluidic channel, a large pressure jump is often used to push
the vesicle through a closely-fit channel~\cite{abk-fai-sto2006}. Our
simulations show that such a configuration can lead to an amplification
of water permeation at a fast time scale of the order of milliseconds.
In addition, we find that a large value of $\beta$ is needed to keep the
membrane tension below the threshold value for membrane poration. Such
high permeability for water may be regarded as an indication that the
lipid bilayer membrane is porated under strong confinement, and our
inextensible membrane model is for a porated vesicle within the Helfrich
free energy framework. As in the previous examples, we fix the membrane
length and examine the enclosed water content as a dynamic consequence
of water flow, membrane shape deformation, and membrane tension
distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Closely-fit Channel (Stenosis)}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semipermeable
  inextensible membrane passing through a closely-fit channel similar to
  the experimental device in~\citet{abk-fai-sto2006}. (b) A
  semipermeable inextensible membrane passing through a (three-times)
  longer closely-fit channel than (a). In (a) and (b), red regions
  correspond to influx and blue regions correspond to efflux. (c) and
  (d) The tension in N/m of the vesicle shapes in (a) and (b). (e) The
  excess pressure of an impermeable and semipermeable inextensible
  membrane in the short geometry in (a). (f) The reduced area in the
  geometries in (a) and (b) as a function of the membranes' center of
  mass. In (e) and (f), the x-axis is scaled by $1/L$ where $L$ is half
  the length of the constriction. (g) The reduced area in the geometries
  in (a) and (b) as a function of time. The time axis is scaled by
  $1/t_R$ where $t_R$ is the residency time. In (e) and (f), the dashed
  lines correspond to the locations of the inlet and outlet. In (g), the
  dashed lines correspond to the times that the vesicle enters and exits
  the constriction.}
\end{figure*}
When red blood cells go through small capillary vessels, they experience
large shape deformation and significant membrane stretching that
triggers ATP release~\cite{Wan2008_PNAS, ForsythWan2011_PNAS}. Many
cells exhibit extraordinary flexibility as they go through narrow
vessels~\cite{AuStoreyMoore2016_PNAS}, and adjustment of
surface-to-volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going
through a stenosis (a long microfluidic channel with a width smaller
than the cell radius). Based on previous
findings~\cite{abk-fai-sto2006}, we first examine the effect of
semipermeability on the pressure fluctuation across a stenosed geometry.
In this experiment, a pressure drop is used to squeeze a red blood cell
through a constriction with diameter $5\;\mu$m and length $45\;\mu$m. We
set up a similar computational geometry with a Poiseuille flow imposed
at the inlet and outlet (Figure~\ref{fig:stenosisComposite}(a)). We set
the maximum flow velocity to $U_{\max} = 1000 \;\mu$m/s at the inlet and
outlet so that the residency time of the membrane in the stenosis
channel is $t_R \approx 25$ ms which agrees with the experimental
results~\cite{abk-fai-sto2006}. When the red blood cell enters the
channel, an increase in the pressure drop is required to maintain a
constant flow rate, and the difference in the pressure drop is called
the excess pressure.

@@

Figures~\ref{fig:stenosisComposite}(f)
and~\ref{fig:stenosisComposite}(g) show that the semipermeable
inextensible membrane deflates from the beginning due to the high flow
velocity, consistent with the results for a vesicle in a Poiseuille flow
in open space. As the membrane enters the stenosis channel, it continues
to deflate in the front even though we observe influx (red segment) at
various rear locations in the membrane. Overall, the membrane loses
about 6.1\% of its initial total water at a constant rate as shown in
Figures~\ref{fig:stenosisComposite}(f)
and~\ref{fig:stenosisComposite}(g) (blue curves). Thus we expect that a
larger amount of deflation is possible if the inextensible membrane is
inside a longer stenosis: we consider a vesicle passing through a
channel of triple the length in Figure~\ref{fig:stenosisComposite}(b)
and (d) with $U_{\max}=1000\;\mu$m/s, and we observe that vesicle stays
inside the stenosis for a residence time $t_R \approx 70$ ms and the
area decreases by 13.2\% (red curves).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Contracting Geometry} 
\begin{figure}[htp]
  \centering
    \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting channel
  similar to the experimental configuration in~\citet{wu2015critical}.
  The channel width varies from $18\;\mu$m to $1.8\;\mu$m, and it
  repeats every $45\;\mu$m. The red regions correspond to influx and
  blue regions correspond to efflux (a--c) The ability for a
  semipermeable vesicle to pass through the contraction depends on the
  permeability $\beta$. (d--f) The ability for a semipermeable
  inextensible membrane to pass through the contraction depends on the
  flow rate. (g--i) The tension distribution and maximum tension in N/m
  of the vesicle shapes in (a) and (b).  (j--l) The tension distribution
  and maximum tension in N/m of the vesicle shapes in (d) and (e).}
\end{figure}

Squeezing a cell or a vesicle through a slit in microfluidics
has gained popularity for its application in stress-induced release of
macromolecules~\cite{ShareiEtAl2013_PNAS, Pak2015_PNAS,
ZhangShenHoganBarakatMisbah2018_BJ, LuoBai2019_PoF}. Red blood cells go
through submicron slits many times during their lifetime, and this
process selects healthy, flexible red blood cells to
survive~\cite{wu2015critical,LuPeng2019_PoF}. Here we investigate the
role of membrane permeability to water when a semipermeable inextensible membrane is
squeezed through a contracting microfluidic channel similar
to~\citet{wu2015critical}, where the width of the channel gradually
decreases from $18 \;\mu$m to $1.8\; \mu$m. The channel then immediately
reopens to $18\;\mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant
maximum flow rate of $U_{\max}$. We initialize a semipermeable inextensible membrane
with $\alpha = 0.65$ that spans $10\; \mu$m of the total channel width
and investigate the effects of the membrane permeability to water and
the maximum flow velocity on whether a membrane can successfully pass
through the slit. 
%We color the vesicle according to the sign of its
%flux---red for influx and blue for efflux.

@@

\begin{figure}[hbp]
  \centering
  \ifTikz
  \input{figures/contractingComposite2.tikz}
  \else
  \includegraphics{figures/contractingComposite2.pdf}
  \fi
  \caption{\label{fig:contractingComposite2} The reduced area and
  velocity of a semipermeable inextensible membrane vs.~time. The
  permeability coefficient is $\beta = 10^{-3}$ and the maximum velocity
  is $18\;\mu$m/s. (a) A semipermeable inextensible membrane passing
  through six periods of the contracting geometry. (b) The reduced area
  with the marks representing shapes in (a). (c) The center-of-mass
  velocity with the marks representing shapes in (a). In (b) and (c),
  the lower $x$-axis is the x-coordinate of the center of mass, and the
  upper $x$-axis is time. (d) The membrane tension in N/m. (e) The
  maximum membrane tension.}
%  The colored vesicles show the tension
%  distribution in N/m of the circled vesicles. Throughout the entire
%  simulation, these vesicle configurations have the largest tension.}
\end{figure}
To understand the effects of water permeability on a semipermeable
inextensible membrane that repeatedly passes through micro-capillary
vessels or sub-micron slits, we construct a contracting geometry
(Figure~\ref{fig:contractingComposite2}) to simulate a membrane passing
through the contracting geometry six times. In
Figure~\ref{fig:contractingComposite2}(b), we show that the membrane
deflates each time it passes through the contraction followed by a
gradual inflation (over seconds). The total area loss over the history
of the vesicle is almost 50\%. In
Figure~\ref{fig:contractingComposite2}(c), we show the membrane
center-of-mass velocity as a function of the membrane position. As the
membrane passes through the contraction, it first accelerates until it
reaches the neck where it abruptly decelerates to go through the
contraction.  Figure~\ref{fig:contractingComposite2}(d) shows the
tension distribution, and the maximum tension is in
Figure~\ref{fig:contractingComposite2}(e). We see that the membrane
tension is largest when it passes through the neck, and it is
particularly large when it passes through the first neck since this is
when the membrane has the largest area.
